Quantitative, Multispecies Monitoring at a Continental Scale

### 2025-02-24 DAG-Ole.tex
Update DAG based on conversations on Friday.

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[node distance=1.5cm]
        % Nodes
\node (qPCR_st) [node, fill=orange!35] at (-4,-2) {
    $\begin{array}{c}
       % \mu_{ik} = \beta_{0p} + \beta_{1p}  K_{i} \\
        \sigma_{Yj} = \exp(\gamma_0 + \gamma_1  K_{j}) \\
        Z_{jpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{j}} \phi) \right) \\
        Y_{jpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{j} , \sigma_{Yj}) \quad \textnormal{if } Z_{jpr} = 1 \\
    \end{array}$
};
\node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standards };
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_st.south east) {};

\node (qPCR_en) [node, below=of qPCR_st, fill=orange!35]  at (0,-3) {
    $\begin{array}{c}
       % \mu_{ip} = \beta_{0p} + \beta_{1p} E_{Hi} \\
        \sigma_{Yi} = \exp(\gamma_0 + \gamma_1 E_{Hi})\\
         Z_{ipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hi}} \phi) \right) \\
         Y_{ipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hi} , \sigma_{Yi}) \quad \textnormal{if } Z_{ipr} = 1
    \end{array}$
};
\node [below=of qPCR_en, yshift=+1.5cm, font=\large\bfseries] {Hake qPCR Likelihood};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_en.south east) {};

\node (state1) [node, above=of qPCR_en, fill=myBlue] at (2,0){ 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
        D_{H}(x,y,d) = \psi  + \epsilon({x,y,d}) \\
        \epsilon({x,y,d})\sim \mathrm{N}(0,\tau)
    \end{array}$
};
\node [above=of state1, yshift=-1.5cm, xshift=0cm, font=\large\bfseries] {Hake DNA (log conc.)};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (state2H) [node, above=of qPCR_en, fill=myBlue] at (2.5,-2.5){ 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
       E_{Hi}(x,y,d) = D_H({x,y,d}) +  \log{V_i} + \\
       		\qquad \qquad \qquad \qquad \log{I_i} + \mathbf{I}\zeta + \delta_i(x,y,d) \\
		\delta_i(x,y,d) \sim \mathrm{N}(0,\sigma_D) \qquad \qquad
    \end{array}$
};
\node [above=of state1, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (state2s) [node, right=of state2H, fill=myBlue]at (4.5,0) { 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
        E_{s \neq H i}(x,y,d) 
    \end{array}$
};
\node [right=of state2s, yshift=0.25cm, xshift=-1.25cm, font=\large\bfseries] {DNA (log conc.) };
\node [right=of state2s, yshift=-0.25cm, xshift=-1.1cm, font=\large\bfseries] {species \textit{s} in bottle \textit{i} };
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (metabarcoding) [node, below=of state2s, fill=orange!35]  at (8,-3) {
    $\begin{array}{c}
        \nu_{si} = (E_{si} - E_{Ri} ) + N_{PCR}  \alpha_s  \\  [8pt]
        \pi_{si} = \frac{exp({\nu_{si}})}{\sum_{s=1}^{S} exp({\nu_{si}})} \\ [8pt]
         \mathbf{R}_{i} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{i}, T_{i} \right) 
    \end{array}$
};
\node [below=of metabarcoding, yshift=1.5cm, xshift=0cm, font=\large\bfseries] {Metabarcoding Likelihood};
\node[anchor=south east] at ([xshift=4pt, yshift=-15pt] metabarcoding.south east) {};

@@

\node (spatial) [node, right=of state2s, fill=green!40] at (8,2){
    $\begin{array}{c}
        F_s(x,y,d) = s\left(E_{s\neq H}(x,y,d) \right)
    \end{array}$
};
\node [above=of spatial, yshift=-1.50cm, xshift=0cm, font=\large\bfseries] {Smooth DNA predictions (conc.) };
\node[anchor=south east] at ([xshift=19pt, yshift=-4pt] spatial.south east) {};

        % Arrows
        \draw [arrow] (qPCR_st) -- node[pos=0.1, below] {
        $\begin{array}{c}
            \phi, \\ \beta_0, \beta_1, \\ \gamma_0, \gamma_1
        \end{array}$} (qPCR_en);
        \draw [arrow] (state1) -- node[midway, left] {$$} (state2H);
        \draw [arrow] (state2H) -- node[midway, right] {} (qPCR_en);
        \draw [arrow] (state2s) -- node[midway, right] {}  (spatial);
        %\draw [arrow] (state1) -- node[midway, right] {}  (spatial);
        \draw [arrow] (mock) -- node[pos=0.7, above] {$\alpha_s$} (metabarcoding);
        \draw [arrow] (state2H) -- node[midway, right] {$$} (metabarcoding);
        \draw [arrow] (state2s) -- node[midway, right] {$$} (metabarcoding);

@@

% Legend: Subscripts
\node (subscripts) [rectangle, draw, align=left, below=of qPCR_en, text width=6cm, fill=myGrey] at (-4.5,-6) {
    \textbf{Index} \\
    $s$ = species ($H$ = Hake) \\
    $x$ = northing \\
    $y$ = easting \\
    $d$ = depth \\
    $i$ = biological replicate \\
    $r$ = technical replicate \\
    $p$ = qPCR plate \\
    $m$ = mock community sample \\
    $j$ = qPCR standard aliquote sample \\
    $R$ = metabarcoding reference species 
};

>>>>

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[node distance=1.5cm]
        % Nodes
\node (qPCR_st) [node, fill=orange!35] at (-4,-2) {
    $\begin{array}{c}
       % \mu_{ik} = \beta_{0p} + \beta_{1p}  K_{i} \\
        Z_{qpr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{K_{q}} \phi) \right) \\
        Y_{qpr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p}  K_{q} , \sigma_{Yq}) \quad \textnormal{if } Z_{qpr} = 1 \\
        \sigma_{Yq} = \exp(\gamma_0 + \gamma_1  K_{q}) \\
    \end{array}$
};
\node [above=of qPCR_st, yshift=-1.5cm, font=\large\bfseries] {qPCR Standards };
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_st.south east) {};

\node (qPCR_en) [node, below=of qPCR_st, fill=orange!35]  at (0,-3) {
    $\begin{array}{c}
       % \mu_{ip} = \beta_{0p} + \beta_{1p} E_{Hi} \\
         Z_{ipr} \sim \mathrm{Bernoulli} \left( 1 - \exp(-2 e^{E_{Hji}} \phi) \right) \\
         Y_{ipr} \sim \mathrm{N} (\beta_{0p} + \beta_{1p} E_{Hji} , \sigma_{Yi}) \quad \textnormal{if } Z_{ipr} = 1 \\
         \sigma_{Yi} = \exp(\gamma_0 + \gamma_1 E_{Hi})
    \end{array}$
};
\node [below=of qPCR_en, yshift=+1.5cm, font=\large\bfseries] {Hake qPCR Likelihood};
\node[anchor=south east] at ([xshift=5pt, yshift=-15pt] qPCR_en.south east) {};

\node (state1) [node, above=of qPCR_en, fill=myBlue] at (2,0){ 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
        D_{Hj} = \psi  + \epsilon_j \\
        \epsilon_j \sim \mathrm{N}(0,\tau)
    \end{array}$
};
\node [above=of state1, yshift=-1.5cm, xshift=0cm, font=\large\bfseries] {Hake DNA (log conc.)};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (state2H) [node, above=of qPCR_en, fill=myBlue] at (2.5,-2){ 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
       E_{Hji} = D_{Hj} +  \log{V_i} +\log{I_i} + \mathbf{I}\zeta + \delta_{ji} \\
       		%\qquad \qquad \qquad \qquad 
		\delta_{ji} \sim \mathrm{N}(0,\sigma_D) \qquad \qquad
    \end{array}$
};
\node [above=of state1, yshift=-1.5cm, xshift=-1.3cm, font=\large\bfseries] {};
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (state2s) [node, right=of state2H, fill=myBlue]at (4.5,0.1) { 
% + \delta0_{ijdbt} 
% + \delta1_{ijdb}
    $\begin{array}{c}
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + \omega_{ijdbt} + \xi_{ijdbt} \\[8pt]
        %\ln C_{ijdbt} = \psi_i + \ln D_{ijd} + \rho_{ijdbt} + W_{ijdbt} + \xi_{ijdbt} \\[8pt]
        E_{s \neq Hji}
    \end{array}$
};
\node [right=of state2s, yshift=0.25cm, xshift=-1.25cm, font=\large\bfseries] {DNA (log conc.) };
\node [right=of state2s, yshift=-0.25cm, xshift=-1.1cm, font=\large\bfseries] {species \textit{s} in bottle \textit{i} };
\node[anchor=south east] at ([xshift=2pt, yshift=-15pt]state1.south east) {};

\node (metabarcoding) [node, below=of state2s, fill=orange!35]  at (8,-3) {
    $\begin{array}{c}
        \nu_{si} = (E_{sji} - E_{Rji} ) + N_{PCR}  \alpha_s  \\  [8pt]
        \pi_{si} = \frac{exp({\nu_{si}})}{\sum_{s=1}^{S} exp({\nu_{si}})} \\ [8pt]
         \mathbf{R}_{i} \sim \mathrm{Multinomial} \left( \boldsymbol{\pi}_{i}, T_{i} \right) 
    \end{array}$
};
\node [below=of metabarcoding, yshift=1.5cm, xshift=0cm, font=\large\bfseries] {Metabarcoding Likelihood};
\node[anchor=south east] at ([xshift=4pt, yshift=-15pt] metabarcoding.south east) {};

@@

\node (spatial) [node, right=of state2s, fill=green!40] at (9,2){
    $\begin{array}{c}
        F_s(x,y,d) 
    \end{array}$
};
\node [above=of spatial, yshift=-1.50cm, xshift=0cm, font=\large\bfseries] {Smooth DNA predictions (conc.) };
\node[anchor=south east] at ([xshift=19pt, yshift=-4pt] spatial.south east) {};

        % Arrows
        \draw [arrow] (qPCR_st) -- node[pos=0.1, below] {} (qPCR_en);
        % \begin{array}{c}
        %     \phi, \\ \beta_0, \beta_1, \\ \gamma_0, \gamma_1
        % \end{array}$} (qPCR_en);
        \draw [arrow] (state1) -- node[midway, left] {$$} (state2H);
        \draw [arrow] (state2H) -- node[midway, right] {} (qPCR_en);
        \draw [dashed,arrow] (state2s) -- node[midway, right] {}  (spatial);
        %\draw [arrow] (state1) -- node[midway, right] {}  (spatial);
        \draw [arrow] (mock) -- node[pos=0.7, above] {$\alpha_s$} (metabarcoding);
        \draw [arrow] (state2H) -- node[midway, right] {$$} (metabarcoding);
        \draw [arrow] (state2s) -- node[midway, right] {$$} (metabarcoding);

@@

% Legend: Subscripts
\node (subscripts) [rectangle, draw, align=left, below=of qPCR_en, text width=6cm, fill=myGrey] at (-4.5,-6) {
    \textbf{Index} \\
    $s$ = species ($H$ = Hake) \\
    $j$ = location (northing, easting, depth) \\
    $x$ = northing \\
    $y$ = easting \\
    $d$ = depth \\
    $i$ = biological replicate \\
    $r$ = technical replicate \\
    $p$ = qPCR plate \\
    $m$ = mock community sample \\
    $q$ = qPCR standard aliquote sample \\
    $R$ = metabarcoding reference species 
};
